### 2015-10-01 analysis/essential.tex

%%%%%%%%%%%%% Summary table of (use, agree) >= (1, 2)
$14$ & \texttt{Class, Component, UseCase, Variant} & \texttt{Comment, Example, Max, Min, Title} & \texttt{implements, verifies}\\\hline
$13$ & \texttt{Configuration, Data, Design, Event, Quality, Scenario, Stakeholder, System, Term} & \texttt{Code, Constraints, Cost, FileName, Probability, Profit, Spec, Why} & \texttt{excludes, interactsWith, is, relatesTo, requires}\\\hline
$12$ & \texttt{Actor, Domain, Feature, Function, Interface, Module, Relationship, Release, Req, Risk, Service, State, Task, Test} & \texttt{Benefit, Capacity, Frequency, Input, Order, Output, Prio, Text, Value} & \texttt{has, impacts}\\\hline
$11$ & \texttt{Idea, Label, Member, Meta, MockUp, Section, User} & \texttt{Image} & \texttt{precedes, superOf}\\\hline
$10$ & \texttt{Goal, Story} & \texttt{Expectation} & \texttt{}\\\hline
$9$ & \texttt{App, Issue, Target, WorkPackage} & \texttt{Damage} & \texttt{binds, helps}\\\hline
$8$ & \texttt{Item, Product, Resource, VariationPoint} & \texttt{} & \texttt{deprecates}\\\hline
$7$ & \texttt{Breakpoint, Screen} & \texttt{Status} & \texttt{}\\\hline
$6$ & \texttt{Barrier} & \texttt{Deprecated} & \texttt{hurts}\\\hline
$5$ & \texttt{} & \texttt{} & \texttt{}\\\hline
$4$ & \texttt{Ticket} & \texttt{} & \texttt{}\\\hline
$3$ & \texttt{} & \texttt{} & \texttt{}\\\hline
$2$ & \texttt{} & \texttt{} & \texttt{}\\\hline
$1$ & \texttt{Epic} & \texttt{Gist} & \texttt{}\\\hline
$0$ & \texttt{} & \texttt{} & \texttt{}\\\hline