### 2015-10-01 analysis/relation-defs.tex

\texttt{binds}&Ties a value to an option. A configuration binds a variation point.\\
\texttt{deprecates}&Makes outdated. An entity deprecates (supersedes) another entity.\\
\texttt{excludes}&Prevents a combination. An entity excludes another entity.\\
\texttt{has}&Expresses containment, substructure. An entity contains another entity.\\
\texttt{helps}&Positive influence. A goal helps to fulfil another goal.\\
\texttt{hurts}&Negative influence. A goal hinders another goal.\\
\texttt{impacts}&Some influence. A new feature impacts an existing component.\\
\texttt{implements}&Realisation of. A module implements a feature.\\
\texttt{interactsWith}&Communication. A user interacts with an interface.\\
\texttt{is}&Sub-typing, specialization, part of another, more general entity.\\
\texttt{precedes}&Temporal ordering. A feature precedes (is implemented before) another feature.\\
\texttt{relatesTo}&General relation. An entity is related to another entity.\\
\texttt{requires}&Requested combination. An entity is required (or wished) by another entity.\\
\texttt{superOf}&Super-typing, generalization, includes another, more specific entity.\\
\texttt{verifies}&Gives evidence of correctness. A test verifies the implementation of a feature.\\