### 2015-10-01 paper/paper-refsq2016-submission.tex

%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is the LaTeX source for the instructions to authors using
% the LaTeX document class 'llncs.cls' for contributions to
% the Lecture Notes in Computer Sciences series.
% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
%
% It may be used as a template for your own input - copy it
% to a new file with a new name and use it as the basis
% for your article.
%
% NB: the document class 'llncs' has its own and detailed documentation, see
% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\modelname, \modelname

%\documentclass[runningheads,a4paper]{llncs}
\documentclass[a4paper]{llncs}

\usepackage{tikz}
\usetikzlibrary{positioning,fit}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
%%%%TODO: REMOVE THE TERM metacompleteness
%% MY COOL PACKAGES
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
%\usepackage{times}
\usepackage{tgtermes}
%\usepackage{bera}
\usepackage[scaled=0.93]{beramono}
\usepackage[normalem]{ulem}
\usepackage{cite}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{scalefnt}
\usepackage{setspace}
%
%\usepackage{listings}
%\lstset{basicstyle=\ttfamily\small, 
%%  numbers=left, numberstyle=\tiny, stepnumber=1, numbersep=5pt,
%  language=java, mathescape=true, escapechar=\!,
%  emph={object}, emphstyle=\textbf,
%  showstringspaces=false
%}
%%%%

\usepackage{hyperref}
\usepackage{xcolor}
\definecolor{entityColor}{RGB}{0,100,200}
\definecolor{attributeColor}{RGB}{0,100,50}
\definecolor{relationColor}{RGB}{160,0,30}
\usepackage{listings}
\lstdefinestyle{reqT}{
  %belowcaptionskip=1\baselineskip,
  breaklines=true,
  %showstringspaces=false,
  showspaces=false,
  %breakatwhitespace=true,
  basicstyle=\ttfamily\scriptsize,
  emph={Ent,Meta,Item,Label,Section,Term,Actor,App,Component,Domain,Module,Product,Release,Resource,Risk,Service,Stakeholder,System,User,Class,Data,Input,Member,Output,Relationship,Design,Screen,MockUp,Function,Interface,State,Event,Epic,Feature,Goal,Idea,Issue,Req,Ticket,WorkPackage,Breakpoint,Barrier,Quality,Target,Scenario,Task,Test,Story,UseCase,VariationPoint,Variant},
  emphstyle=\bfseries\color{entityColor},
  emph={[2]has,is,superOf,binds,deprecates,excludes,helps,hurts,impacts,implements,interactsWith,precedes,requires,relatesTo,verifies},
  emphstyle={[2]\color{relationColor}},
  emph={[3]Attr,Code,Constraints,Comment,Deprecated,Example,Expectation,FileName,Gist,Image,Spec,Text,Title,Why,Benefit,Capacity,Cost,Damage,Frequency,Min,Max,Order,Prio,Probability,Profit,Value,Status},
  emphstyle={[3]\itshape \color{attributeColor}},  
}
\lstset{style=reqT}

\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}
\begin{document}

%\floatstyle{ruled}
%\newfloat{Specification}{htbp}{lop}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{What is essential? -- A pilot survey on views about the requirements metamodel of reqT}
%reqT --  A Requirements Modeling Tool for Code Lovers
% a short form should be given in case it is too long for the running head
%\titlerunning{DRAFT MANUSCRIPT - Please do not cite. Submitted to REFSQ'13.}

\author{Bj\"orn Regnell}
%
%\authorrunning{Regnell, "Requirements Modeling for Code Lovers"}
\institute{Dept. of Computer Science, Lund University, Sweden \\ \url{bjorn.regnell@cs.lth.se} }

\maketitle

%%%% ABSTRACT
\begin{abstract}
[{\bf Context \& motivation}] This research preview presents ongoing work on the metamodel of a free software requirements modeling tool called reqT that is developed in an educational context. The work aims to make an initial validation of a survey instrument that elicits views on the metamodel of the reqT tool, which seek to engage computer science students in Requirements Engineering (RE) through an open source requirements engineering DSL embedded in the Scala programming language. [{\bf Question}] The research question is: Which RE concepts are essential to include in the metamodel for a requirements engineering tool in an educational context?  [{\bf Principal ideas}] A survey instrument is developed with a list of 92 concepts (49 entities, 15 relations and 28 attributes) and a set of questions for each concept that elicit the respondents' views on the usage and interpretation of each concept.  [{\bf Contribution}] The survey is initially validated in a pilot study involving 14 Swedish RE scholars as subjects. The survey results indicate that the survey is feasible if the respondents is willing to invest around 30 minutes of their time. The analysis of the responses suggest that many of the concepts in the metamodel are used frequently by the respondents and there is a large degree of agreement among the respondents about the meaning of the concepts. Some terms can be viewed as ''essential RE concepts'' in that a many use them and agree on their meaning. The results are encouraging for future work on empirical validation of the relevance of the reqT metamodel.

\keywords{requirements engineering, requirements metamodel, CASE tool, requirements engineering education, embedded domain-specific language, empirical software engineering}
\end{abstract}

%%%%% INTRO
\section{Introduction}
There are many challenges in teaching Requirements Engineering (RE) \cite{Memon2010, Regev2011}, including  conveying requirements modelling skills that can be used effectively in an unstructured, non-ideal, real-world situation \cite{Callele2006}. When teaching RE modelling we may ask ourselves: What are the \textit{essential} RE concepts that we should include in our taught metamodel for requirements? This paper investigates this questions in conjunction with the on-going work of developing a metamodel for reqT \cite{reqT}, an open source requirements engineering tool \cite{Regnell2013} used in RE education \cite{ets170}.
A survey instrument is presented aiming to elicit the frequency of RE term usage and the degree of interpretation agreement. The responses from 15 Swedish RE scholars are analysed and discussed and conclusions suggest that a large subset of the concepts of the current reqT metamodel can be argued to be ''essential'' in that a majority of the subjects use them while agreeing with the concept definitions. The presented work is an initial validation and further work involving more subjects is needed to provide conclusions with more certainty.

\section{Background and Related Work}

There are nowadays numerous commercial RE tools available, but many are expensive, complex and not sufficiently open  \cite{Carillo2011}. A major aim of the reqT open source project is to provide a small but scalable, semi-formal and free software package for an educational setting \cite{Regnell2013} that can inspire code-loving computer science students to learn more about requirements modeling. The tool development started in 2011 at Lund University, where reqT is used in RE teaching at MSc level in the Computer Science \& Engineering program \cite{ets170}.

A critical issue is how to find the "essential" RE concepts that allows for sufficient expressiveness, while not overloading the metamodel with esoteric concepts just for the sake of completeness. 
%The reqT tool is used in a course based on a specific text book \cite{Lauesen2002} and a specific student project concept \cite{ets170}, and the concepts of the earlier versions of the reqT requirements metamodel reflect that context.

The reqT metamodel includes three types of concepts: entities, attributes and relations. Entities and attributes are nodes in a graph data structure, while relations are edges that can connect entities with sub-graphs. Thus a tree-like structure can be created of arbitrary depth spanning the graph that models some chunk of requirements.

Version 3.0 of the metamodel includes the concepts defined in Appendix A. These concepts and definitions are gatherd from various sources including the IREB Glossary \cite{IREB}, wikipedia, terminology from agile development, variability modelling terminology, and the Lauesen text book \cite{Lauesen2002} used in the RE course at Lund University \cite{ets170}, in which reqT is applied in student role-playing projects.

!!! EXPLAIN EXAMPLE !!!
\begin{spacing}{0.8}
\begingroup
    \fontsize{8pt}{12pt}\selectfont
\begin{lstlisting}
Model(
  Component("apperance") has (
    VariationPoint("color") has (
      Min(0), Max(2),
      Variant("blue"), Variant("red"), Variant("green")),
    VariationPoint("shape") has (
      Min(1), Max(1), Variant("round"), Variant("square")),
    VariationPoint("payment") has (
      Min(1), Max(2), Variant("cash"), Variant("credit")),
    VariationPoint("payment") requires Variant("cash"), 
    Variant("round") excludes Variant("red"),
    Variant("green") requires Variant("square")),
  Component("apperance") requires VariationPoint("shape"), 
  App("free") requires Component("apperance"),  
  App("free") binds (
    VariationPoint("shape") binds Variant("round")),
  App("premium") requires Component("apperance"),  
  App("premium") binds ( 
    VariationPoint("color") binds (Variant("red"), Variant("green")),
    VariationPoint("shape") binds (Variant("round"), Variant("square")),
    VariationPoint("payment") binds Variant("cash")))
\end{lstlisting}
\endgroup
\end{spacing}

\section{Methodology and Data Collection}
!!! explain how sources such as the IREB glossary was used to form the metamodel concept definitions !!!
\section{Data Analysis}

\textbf{Subject background.} The background questions in the survey regards the role of the subject, as shown in Table \ref{table:background}. In summary, the included\footnote{One subject answered NO on all background questions and was therefore excluded.} \input{../analysis/summary.tex}
\begin{table}[H]
\centering
\fontsize{8}{9}\selectfont
\caption{Background of subjects, $N = 15$. The table shows anonymized subject ids}
\label{table:background}
\begin{tabular}{p{0.42\textwidth}| p{0.58\textwidth}}
\textit{Question} & \textit{Subject responding YES}  \\ \hline
\input{../analysis/background.tex}

 \end{tabular}
\end{table}

\textbf{Frequency analysis.} The idea of ''essentiality'' is characterized as the number of subjects that has responded that they (1) use the concepts at least in an informal, non-persistent way, and (2) use the concepts in a similar meaning as in the definition in Appendix A. Table \ref{table:frequency} shows the results of the frequency counts.

%
\begingroup
\setlength{\tabcolsep}{4pt} % Default value: 6pt
\renewcommand{\arraystretch}{1.5} % Default value: 1
\begin{table}[H]
\centering
\fontsize{7}{8}\selectfont
\caption{Frequency analysis of concepts' ''essentiality'', where $n$ is the number of subjects that answered $(use >= 1)$ and $(agree = 2)$ for each concept.}
\label{table:frequency}
\begin{tabular}{l | p{0.33\textwidth} | p{0.30\textwidth} | p{0.25\textwidth}}
\textit{$n$} & \textit{Entities} & \textit{Attributes} & \textit{Relations} \\ \hline
\input{../analysis/essential.tex}
 \end{tabular}
\end{table}
\endgroup
%%%%%%%%%%%%%% DISCUSSION

\section{Discussion and Conclusion}\label{section:discussion}

!!Discuss and conclude!!

{\bf Limitations. } Due to the limited number of subjects and the high degree of homogeneity among subjects with respect to background, it is difficult to analyse and draw conclusions about potential differences in opinions between e.g. teachers and developers. Some subjects needed more time and completed survey offline during the coming days, which may give a variation in how careful the responses were considered. The survey was conducted by the author and inventor of reqT, in conjunction with a seminar and demo of reqT. In order to avoid any positive bias due to advocacy in favour of the reqT metamodel, the survey was held prior to the seminar and demo. This in turn may introduce a threat of limited knowledge among subjects of the idea behind the modelling approach in reqT

{\bf Future work.} %Further directions of research include (1) incorporation of constraints on models for support of prioritization and release planning \cite{Regnell2011}, (2) more elaborate semantic checks to better guide requirements modelers, and (3) graphical visualization of requirements graph models. (4) Natural Language Processing technology including e.g. ambiguity risk detection may be interesting to combine with reqT. (5) It is also important to further investigate the pedagogic advantages and limitations of the approach.

%A major objective of this research preview paper is to expose the latest version of reqT to the community of RE scholars and to invite discussions and contributions.

\subsubsection*{Acknowledgments.} This work is partly funded by  VINNOVA within the EASE project.

\bibliography{refs}
\bibliographystyle{splncs03}

%%%%%%%%%%%%%%% APPENDIX WITH DEFINITIONS
\clearpage
%\hspace{-0.9cm}
\section*{Appendix A: Definitions of Metamodel Concepts}
\begin{minipage}{1\linewidth}
\begin{minipage}{0.30\textwidth}

\begin{table}[H]
%\centering
\vspace{-0.8cm}
\fontsize{6}{6.5}\selectfont
%\caption{.}%: Entities, Attributes and Relations.}

\begin{tabular}{|l p{4.5cm}|}
\hline
\textit{Entity} & \textit{Definition}  \\ \hline
\input{../analysis/entity-defs.tex}
\hline
 \end{tabular}
\end{table}
\end{minipage}
 \hspace{4em plus 1fill} 
\begin{minipage}{0.47\textwidth} 
 \begin{table}[H]
%\centering
\fontsize{6}{6.5}\selectfont
%\caption{Definitions of attributes and relations.}
\vspace{-0.8cm}
\label{table-entities}
\begin{tabular}{|l p{4.0cm}|}
 \hline
\textit{Attribute} & \textit{Definition}  \\ \hline
\input{../analysis/attribute-defs.tex}

\hline
\textit{Relation} & \textit{Definition}  \\ \hline
\input{../analysis/relation-defs.tex}
 \hline
\end{tabular}
\end{table}

\end{minipage}
\end{minipage}

\end{document}