### 2016-03-14 paper/proceedings/attribute-defs.tex
rename dirs

\texttt{Benefit}&A characterisation of a good or helpful result or effect (e.g. of a feature).\\
\texttt{Capacity}&The largest amount that can be held or contained (e.g. by a resource).\\
\texttt{Code}&A collection of (textual) computer instructions in some programming language, e.g. Scala. Short for source code.\\
\texttt{Comment}&A note that explains or discusses some entity.\\
\texttt{Constraints}&A collection of propositions that restrict the possible values of a set of variables.\\
\texttt{Cost}&The expenditure of something, such as time or effort, necessary for the implementation of an entity.\\
\texttt{Damage}&A characterisation of the negative consequences if some entity (e.g. a risk) occurs.\\
\texttt{Deprecated}&A description of why an entity should be avoided, often because it is superseded by another entity, as indicated by a 'deprecates' relation.\\
\texttt{Example}&A note that illustrates some entity by a  typical instance.\\
\texttt{Expectation}&The required output of a test in order to be counted as passed.\\
\texttt{FileName}&The name of a storage of serialized, persistent data.\\
\texttt{Frequency}&The rate of occurrence of some entity. \\
\texttt{Gist}&A short and simple description of an entity, e.g. a function or a test.\\
\texttt{Image}&(The name of) a picture of an entity.\\
\texttt{Input}&Data consumed by an entity, \\
\texttt{Max}&The maximum estimated or assigned (relative) value.\\
\texttt{Min}&The minimum estimated or assigned (relative) value.\\
\texttt{Order}&The ordinal number of an entity (1st, 2nd, ...).\\
\texttt{Output}&Data produced by an entity, e.g. a function or a test.\\
\texttt{Prio}&The level of importance of an entity. Short for priority.\\
\texttt{Probability}&The likelihood that something (e.g. a risk) occurs.\\
\texttt{Profit}&The gain or return of some entity, e.g. in monetary terms.\\
\texttt{Spec}&A (detailed) definition of an entity. Short for specification\\
\texttt{Status}&A level of refinement of an entity (e.g. a feature) in the development process. \\
\texttt{Text}&A sequence of words (in natural language).\\
\texttt{Title}&A general or descriptive heading.\\
\texttt{Value}&An amount. An estimate of worth.\\
\texttt{Why}&A description of intention. Rationale.\\