### 2016-03-15 paper/presentation/presentation.tex
add stuff

\documentclass{slides}

\setbeamertemplate{footline}[frame number]
\title[What is essential?]{What is essential?}
\subtitle{\footnotesize -- A pilot survey on views about the requirements metamodel of reqT.org}
\author{Björn Regnell}
\institute{Lund University}
\date{March 14, 2016\\ \href{http://refsq.org/2016}{refsq.org/2016}\\ \href{https://github.com/bjornregnell/reqT-survey}{github.com/bjornregnell/reqT-survey}}

\begin{document}

\frame{\titlepage}
\frame{\tableofcontents}

\section{Objective}
\subsection{Research Question}
\begin{Slide}{Research question}
In the context of software requirements engineering education:
\begin{itemize}
\item How to choose a set of \\ \Emph{essential requirements engineering concepts} \\ that allows for \Alert{sufficient expressiveness}, \\ without overloading the metamodel with esoteric concepts just for the sake of \Alert{completeness}?
\end{itemize}
\end{Slide}
\subsection{Approach}
\begin{Slide}{Approach}
Make a survey among RE scholars:
\begin{itemize}
\item How to quantify ''essentiality''?
\pause
\item One possible quantification: \\ The more scholars that \Emph{agree} on a definition of a concept \\ and \\ the more scholars that \Emph{use} the concept, \\ the more essential is the concept.
\end{itemize}
\end{Slide}

\section{Background}
\subsection{About reqT}
\begin{Slide}{reqT}
\begin{itemize}
\item reqT is an open source requirements engineering tool \\ \url{http://reqt.org/}
\item reqT is developed and used in an \Emph{educational context} aimed at inspiring \Alert{code-loving} computer science students to learn more about RE
\end{itemize}
\end{Slide}

\section{Methodology}
\subsection{Data Collection}
\begin{Slide}{Data Collection}
\end{Slide}

\subsection{Data Analysis}
\begin{Slide}{Data Analysis}
\end{Slide}

\section{Results}
\subsection{Essentiality}
\begin{Slide}{Essentiality}
\end{Slide}

\section{Conclusion}
\subsection{Future Work}
\begin{Slide}{Future Work}
\end{Slide}

\end{document}