### 2016-03-15 paper/presentation/presentation.tex
add more stuff

\documentclass{slides}

\setbeamertemplate{footline}[frame number]
\title[What is essential? -- A research preview]{What is essential?}
\subtitle{\footnotesize -- A pilot survey on views about the requirements metamodel of reqT.org}
\author{Björn Regnell}
\institute{Lund University, Sweden}
\date{\href{http://refsq.org/2016}{refsq.org/2016}\\ \href{https://github.com/bjornregnell/reqT-survey}{github.com/bjornregnell/reqT-survey}}

\usepackage{xcolor}
\definecolor{entityColor}{RGB}{0,100,200}
\definecolor{attributeColor}{RGB}{0,100,50}
\definecolor{relationColor}{RGB}{160,0,30}

\lstdefinestyle{reqT}{
  %belowcaptionskip=1\baselineskip,
  breaklines=true,
  %showstringspaces=false,
  showspaces=false,
  %breakatwhitespace=true,
  basicstyle=\ttfamily\fontsize{8}{10}\selectfont,
  emph={Ent,Meta,Item,Label,Section,Term,Actor,App,Component,Domain,Module,Product,Release,Resource,Risk,Service,Stakeholder,System,User,Class,Data,Input,Member,Output,Relationship,Design,Screen,MockUp,Function,Interface,State,Event,Epic,Feature,Goal,Idea,Issue,Req,Ticket,WorkPackage,Breakpoint,Barrier,Quality,Target,Scenario,Task,Test,Story,UseCase,VariationPoint,Variant},
  emphstyle=\bfseries\color{entityColor},
  emph={[2]has,is,superOf,binds,deprecates,excludes,helps,hurts,impacts,implements,interactsWith,precedes,requires,relatesTo,verifies},
  emphstyle={[2]\color{relationColor}},
  emph={[3]Attr,Code,Constraints,Comment,Deprecated,Example,Expectation,FileName,Gist,Image,Spec,Text,Title,Why,Benefit,Capacity,Cost,Damage,Frequency,Min,Max,Order,Prio,Probability,Profit,Value,Status},
  emphstyle={[3]\itshape \color{attributeColor}},  
}
\lstset{style=reqT, language=}

\begin{document}

\frame{\titlepage}
\frame{\tableofcontents}

\section{Objective}
\subsection{Research Question}
\begin{Slide}{Research question}
In the context of software requirements engineering education:
\begin{itemize}
\item How to choose a set of \\ \Emph{essential requirements engineering concepts} \\ that allows for \Alert{sufficient expressiveness}, \\ without overloading the metamodel with esoteric concepts just for the sake of \Alert{completeness}?
\pause
\item Presumption: as teachers we should be method agnostic; there is no single ''correct'' dogma
\end{itemize}
\end{Slide}
\subsection{Approach}
\begin{Slide}{Approach}
\begin{itemize}

\item Make a survey among RE scholars
\pause
\begin{itemize}
\item How to quantify ''essentiality''?
\pause
\begin{itemize}
\item One possible quantification: \\ The more scholars that \Emph{agree} on a definition of a concept \\ and \\ the more scholars that \Emph{use} the concept, \\ the more \Alert{essential} is the concept.
\end{itemize}
\end{itemize}
\pause
\item Use the reqT.org metamodel as a basis for the survey
\end{itemize}
\end{Slide}

\section{Background}
\subsection{About reqT}
\begin{Slide}{What is reqT?}
\begin{itemize}
\item reqT is an open source requirements engineering tool developed in an \Emph{educational} context\\ Download from here: \url{http://reqt.org/} \\ \texttt{java -jar reqT.jar}
\item reqT aims to inspire \Emph{code-loving} cs students to learn more about RE through a \Emph{Scala}-internal DSL
\item has a metamodel with {\color{entityColor}{\bf entities}}, {\color{relationColor}{\bf relations}} and {\color{attributeColor}{\bf attributes}}\\ gathered from various sources such as \\ text books, research papers, IREB, wikipedia, etc.
\end{itemize}
\end{Slide}

\begin{Slide}{A small reqT model}
\begin{Code}
val m = Model(
  Feature("chat") has (
    Spec("The system shall enable sending text messages among users"),
    Prio(42)
  )
)
\end{Code}
\end{Slide}

\begin{Slide}{An larger example reqT model}
\begin{Code}
Model(
  Component("appearance") has (
    VariationPoint("color") has (
      Min(0), Max(2), Variant("blue"), Variant("red"), Variant("green")),
    VariationPoint("shape") has (
      Min(1), Max(1), Variant("round"), Variant("square")),
    VariationPoint("payment") has (
      Min(1), Max(2), Variant("cash"), Variant("credit")),
    VariationPoint("payment") requires Variant("cash"), 
    Variant("round") excludes Variant("red"),
    Variant("green") requires Variant("square")),
  Component("appearance") requires VariationPoint("shape"), 
  App("free") has Component("appearance"),  
  App("free") binds (VariationPoint("shape") binds Variant("round")),
  App("premium") has Component("appearance"),  
  App("premium") binds ( 
    VariationPoint("color") binds (Variant("red"), Variant("green")),
    VariationPoint("shape") binds (Variant("round"), Variant("square")),
    VariationPoint("payment") binds Variant("cash")))
\end{Code}
\end{Slide}

\begin{Slide}{49 entities of the metamodel (see paper appendix)}
\fontsize{3.5}{4.0}\selectfont
\begin{tabular}{l l}
\textbf{\textit{Entity}} & \textbf{\textit{Definition}}  \\
\input{../proceedings/entity-defs.tex}
\end{tabular}
\end{Slide}

\begin{Slide}{28 atributes of the metamodel (see paper appendix)}
\fontsize{6}{6.5}\selectfont
\begin{tabular}{l l}
\textbf{\textit{Attributes}} & \textbf{\textit{Definition}}  \\
\input{../proceedings/attribute-defs.tex}
\end{tabular}
\end{Slide}

\begin{Slide}{15 realtions of the metamodel (see paper appendix)}
\fontsize{8}{9}\selectfont
\begin{tabular}{l l}
\textbf{\textit{Relation}} & \textbf{\textit{Definition}}  \\
\input{../proceedings/relation-defs.tex}
\end{tabular}
\end{Slide}

\section{Methodology}
\subsection{Data Collection}
\begin{Slide}{Data Collection}
{\centering
\includegraphics[width=0.8\textwidth]{img/survey-screen-dump}
}
{\small Q1 use =   \{no | used, but only orally | used, also in writing\}\\ Q2 agree = \{no, different meaning | don't know | yes, similar meaning\}  \\
Answered by 15 swedish RE scholars (100\% response rate)\\
\url{https://github.com/bjornregnell/reqT-survey}
}

\end{Slide}

\subsection{Data Analysis}
\begin{Slide}{Data Analysis}
\textbf{Frequency analysis.} The degree of ''essentiality'' is characterized as the number of subjects that has responded that they (1) use the concept at least in an informal, non-persistent way, \textit{and} that they (2) use the concept in a similar meaning as in the definition in Appendix A.

\end{Slide}

\section{Results}
\subsection{Essentiality}
\begin{Slide}{Frequency Analysis}
\fontsize{6}{8}\selectfont

\begin{tabular}{l | p{0.33\textwidth} | p{0.30\textwidth} | p{0.25\textwidth}}
\textit{$n$} & \textit{Entities} & \textit{Attributes} & \textit{Relations} \\ \hline
\input{../proceedings/essential.tex}
\end{tabular}

\end{Slide}

\begin{Slide}{Essentiality Definition}
\begin{itemize}
\item If an ''essentiality threshold'' is chosen at $N/2$ then \\ only the $9$ concepts from row $n = 7$ and below are considered ''non-essential''
\item \Alert{More than 90\% of the 92 metamodel concepts have a majority of the subjects that use them and agree upon their definitions} 
\item Each concept has at least one subject that uses it and agrees with its definition. 
\end{itemize}

\end{Slide}

\section{Conclusion}
\subsection{Contribution \& Future Work}
\begin{Slide}{Conlusion}

\begin{itemize} 
\item Contribution
\begin{itemize} 
\item A way to quantify ''essentiality'' -- tested in a pilot survey
\item A set of potentially ''essential'' concepts \\ (according to 14 RE scolars in Sweden)
\end{itemize}
\end{itemize}

\begin{itemize} 
\item Future Work
\begin{itemize} 
\item More subjects for increased external validity
\item More questions on missing and alternative concepts
\end{itemize}
\end{itemize}
\end{Slide}

\end{document}