### 2016-03-17 paper/presentation/discussant.tex
add stuff

\documentclass{slides}

\setbeamertemplate{footline}[frame number]
\title{REFSQ2016}
\begin{document}

\section{Paper discussion in session 5a: RE Foundations}
\subsection{Foundations for Transparency Requirements Engineering, Hosseini et al.}

\begin{Slide}{Discussion items (Björn Regnell)}
\begin{itemize}
\item  \Emph{What can we do now that we could not do before?} \\ \pause
  Apply the framework in our RE practice
\pause\item  \Emph{How sound is the solution?} \\ \pause
  Based on a literature review, ''extensive'', systematic?
\pause\item    \Emph{What is the next step to take?} \\ \pause
        Test it in an empirical case study using real-world data
\pause\item    \Emph{What is the most questionable issue of the paper?} \\ \pause
\begin{itemize}
\item  \Alert{The???} framework?
\item  For what \Alert{treatment} to practical problems is this a cure?
\pause
\begin{itemize}
\item What have you missed? (Completeness)
\item What is essential? Apply ''Occam's razor''
\item What is your moral stand-point on transparency?
\end{itemize}
\end{itemize}
\end{itemize}

\end{Slide}

\end{document}